---
title: "Analysis Phytophthora spp. survey"
author: "Daniel Cerritos"
date: "4/24/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
# install.packages("pacman")
# install.packages(c("googleway", "ggrepel","ggspatial", "libwgeom", "sf", "rnaturalearth", 
# "rnaturalearthdata","hagis", "corrplot", "Hmisc", "drc", "rgeos", "scales"))
pacman::p_load(tidyverse, here, sf, maps, rnaturalearth, rnaturalearthdata, 
               cowplot, hagis, pander, lmerTest, ggResidpanel, emmeans, broom, 
               Hmisc, corrplot, drc, scales)
```


# Import data

```{r data, message=FALSE}
all_data <- list.files(here("output", "clean_data"), full.names = TRUE)
# samples collection
soil_samples <- read_csv(all_data[9])
plant_samples <- read_csv(all_data[8])
# identification of isolates
isolates_recovered <- read_csv(here("data", "all-isolates-in-data-base.csv"))
# pathotype
sojae <- read_csv(all_data[7])
sansomeana <- read_csv(all_data[6])
#aggressiveness
dry_weigth <- read_csv(all_data[1])
area_length <- read_csv(all_data[2])
agg_temperature <- read_csv(all_data[3]) 
# fungicide sensitivity 
fungicide <- read.csv(all_data[5])
```


# Sample collection

```{r soil samples}
# 2016-2017
soil_samples %>%
  group_by(year) %>% 
  count() 

soil_samples$county %>%
  n_distinct()
```

add new file: "Plant_Clinic_2018" to see the samples that were actually processed, only 52 from the 95 that Phytophthotra was detected by the ELISA

```{r plant samples}
# 2018
plant_samples %>%
  count() # just included positive samples 
  
plant_samples$county %>%
  n_distinct()
```

```{r samples data map}
# subset county data 
world <- ne_countries(scale = "medium", returnclass = "sf")
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))
counties <- subset(counties, grepl("illinois", counties$ID))

# convert to sf object
map_soil <- soil_samples %>% 
  filter(!sample_id == "16PR012") %>% # not from IL
  dplyr::select(year, long, lat) %>% 
  filter(!is.na(long))

map_soil <- st_as_sf(map_soil, coords = c("long", "lat"), 
                   crs = 4326, agr = "constant")

# change continuous data (plant samples) into categories 
map_plant <- plant_samples %>% 
  group_by(county) %>% 
  count() %>% 
  mutate(samples_categ = case_when(
    n <= 5 ~ "1-5", 
    n >= 6 & n <= 10 ~ "6-10",
    n >= 11 & n <= 15 ~ "11-15", 
    n >= 16 & n <= 20 ~ "16-20", 
  ))

# add counties coordinates to plant samples data
counties <- counties %>% 
  separate(ID, into = c("state", "county"), sep = ",")

map_plant$county <- tolower(map_plant$county)
map_plant <- left_join(counties, map_plant, by = "county")

# as factor for plotting
map_plant$samples_categ <- as.factor(map_plant$samples_categ)
map_plant$samples_categ <- factor(map_plant$samples_categ, 
                                  levels = c("1-5", 
                                             "6-10",                                  
                                             "11-15",
                                             "16-20" ))
map_soil$year <- as.factor(map_soil$year)
map_plant <- map_plant %>% 
  filter(!is.na(samples_categ))
```

## Figure 1

change this figure to actual 52 samples 

```{r sample map}
# samples collection map 
ggplot() +
  geom_sf(data = counties, fill = NA) +
  geom_sf(data = map_plant, aes(fill = samples_categ) , alpha = 0.7) +
  geom_sf(data = map_soil, aes(shape = year),  size = 2) +
  coord_sf(xlim = c(-92, -87), ylim = c(36.5, 43), expand = FALSE) +
  scale_fill_manual(values = c("#bdd7e7", "#6baed6", "#3182bd", "#08519c"), 
                    name = "Plant samples") +
  scale_shape_discrete(name = "Year") +
  theme_map()
```

# Identification of isolates

```{r exp isolates}
isolates_recovered %>% 
  group_by(identification) %>% 
  count() 

isolates_recovered %>% 
  group_by(source) %>% 
  count()

isolates_recovered %>% 
  filter(!is.na(county)) %>% 
  group_by(year, county) %>% 
  count() %>% 
  ggplot(aes(x = fct_reorder(county, n), y = n, fill = as.factor(year))) +
  geom_col(alpha = 0.6) +
  coord_flip() +
  labs(x = "", y = "Isolates recovered", fill ="Year") +
  theme_minimal_vgrid() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
```

## Figure 2 

```{r species collected}
isolates_species <- isolates_recovered %>% 
  dplyr::select(species) %>%
  group_by(species) %>% 
  count() %>%  
  mutate(genus = case_when(
    grepl("Phyto", species) ~ "Phytophthora spp.",
    grepl("Pyth", species) ~"Pythium spp.")) %>% 
  mutate(percentage = (n/180))

isolates_species %>% 
  ggplot(aes(x = reorder(species, percentage), y = percentage, fill = genus)) +
  geom_col(alpha = 0.6) +
  coord_flip() +
  labs(x = "", y = "Percent isolates recovered") +
  theme_minimal_vgrid() +
  theme(legend.position = "none",
        axis.text.y  = element_text(face = "italic")) +
  scale_fill_manual(values = c("#F8766D", "#BEBEBE")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                     labels = percent_format()) 
```

```{r phytophthora county}
isolates_recovered %>% 
  filter(species %in% c("Phytophthora sansomeana", "Phytophthora sojae")) %>% 
  group_by(year, county) %>% 
  count(species) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```

# Pathotype

## Phytophthora sojae

```{r effective Rps}
# two susceptible controls
sojae_control <- sojae %>% 
  filter(differential == "Williams") %>% 
  group_by(isolate_number, isolate_id, isolate_name, county, differential, gene) %>% 
  summarise(dead = sum(dead), 
            total = sum(total))

sojae <- sojae %>% 
  filter(!differential == "Williams")

sojae <- bind_rows(sojae_control, sojae)

sojae <- sojae %>% 
  mutate(percent.susceptible = (dead/total)*100)

hagis_args <- list(
  x = sojae,
  cutoff = 75,
  control = "rps",
  sample = "isolate_number",
  gene = "gene",
  perc_susc = "percent.susceptible"
)

# effective Rps Fig 3A
sojae_rps <- do.call(summarize_gene, hagis_args)

effective_rps <- sojae_rps %>% 
  mutate(percent_pathogenic = percent_pathogenic/100) %>% # to use the scales format
  ggplot(aes(x = gene, y = percent_pathogenic)) +
  geom_col(alpha = 0.6, fill = "#F37B59") +
  labs(x = "Rps gene", 
       y = "Percent pathogenic") +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)), 
    labels = percent_format()) +
    theme_minimal_hgrid()
```

```{r complexity}
complexities <- do.call(calculate_complexities, hagis_args)

# mean complexity
pander(summary(complexities))

# Fig 3B
complexity_percent <- complexities$grouped_complexities %>% 
  mutate(percent = (frequency/21))

complexity_percent$complexity <- as.numeric(complexity_percent$complexity)

mean_complexity <- complexity_percent %>% 
  ggplot(aes(x = complexity, y = percent)) +
  geom_col(alpha = 0.6, fill = "#529EFF") + 
  geom_vline(xintercept = 6, linetype="dashed",  size = 1) +
  labs(x = "Complexity", 
       y = "Percent Isolates") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                     labels = percent_format()) +
  scale_x_continuous(limits = c(1, 12), breaks = 1:12) +
  theme_minimal_hgrid() 
```

## Figure 3

```{r fig 3, fig.height=5.7, fig.width=4.5, warning=FALSE}
plot_grid(effective_rps, mean_complexity, labels = "AUTO", ncol = 1)
```

## Table 1

```{r pathotype}
# pathotypes
diversity <- do.call(calculate_diversities, hagis_args)
diversities_table(diversity) 

pathotype_frequency <- diversity$table_of_pathotypes
# write_csv(pathotype_frequency, here("output", "results", "02_pathotype-psojae_frequency-table.csv"))
```

## Phytophthora sansomeana

```{r sansomena effective genes, fig.height=2.85, fig.width=4.5}
sansomeana <- sansomeana %>% 
  mutate(percent.susceptible = round((dead/total)*100))

sansomeana_rps <- summarize_gene(
  x = sansomeana,
  cutoff = 75,
  control = "rps",
  sample = "isolate_id",
  gene = "gene",
  perc_susc = "percent.susceptible"
)

sansomeana_rps %>% 
  ggplot(aes(x = gene, y = percent_pathogenic)) +
  geom_col(alpha = 0.5) +
  theme_minimal_hgrid() +
  labs(x = "Rps gene", 
       y = "Percent pathogenic") +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))) 
```

```{r save pathotype data}
# combine pathtoyping results
isolates_complexity <- complexities$indvidual_complexities %>% 
  rename(isolate_number = sample, complexity = N_samp)

isolates_pathotypes <- diversity$individual_pathotypes %>% 
  rename(isolate_number = Sample)

isolates_complexity$isolate_number <- as.numeric(isolates_complexity$isolate_number)
sojae_all <- left_join(isolates_complexity, isolates_pathotypes)

add_isolate_id <- sojae[1:24, ]

sojae_pathotype <- left_join(sojae_all, add_isolate_id, by = "isolate_number") %>% 
  dplyr::select(isolate_number, isolate_id, complexity, Pathotype)

# write_csv(sojae_pathotype, here("output", "results", "02_pathotype-psojae_results.csv"))
```

# Aggressiveness

```{r area and length}
# some pots had two images in experiment 1 because they did not fit
area_length <- area_length %>% 
  group_by(experiment, pot) %>% 
  summarise(length_in = mean(length_in), 
            area_in.2 = mean(area_in.2))

aggressiveness <- left_join(dry_weigth, area_length, 
                            by = c("experiment", "pot")) 

# results  are by pot, divide by germination to get mean by seedling
aggressiveness <- aggressiveness %>% 
  mutate(shoot_mg = shoot.mg/germination, 
         root_mg = root.mg/germination, 
         length_in = length_in/germination, 
         area_in.2 = area_in.2/germination)

# write_csv(aggressiveness, here("output", "transform_data", "02_aggressiveness_transform.csv"))
```

## Exploratory plots 

```{r average temperature, message=FALSE}
agg_temperature %>% 
  group_by(experiment) %>% 
  summarise(min = min(room_temperature), 
            max = max(room_temperature), 
            mean = mean(room_temperature))
```

```{r temperature agg}
agg_temperature$time <- as.character(agg_temperature$time)
# filter to the hours they have in common 
agg_temperature_filtered <- agg_temperature %>% 
  filter(time %in% c("00:15:00", "08:15:00", "16:15:00"))

agg_temperature_filtered %>% 
  ggplot(aes(x = date, y = room_temperature, color = as.factor(experiment))) +
  geom_line(group = 1)  +
  facet_grid(~ as.factor(experiment), scales = "free_x") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "none") 
```

```{r exp plot 1, warning=FALSE}
aggressiveness$experiment <- as.factor(aggressiveness$experiment)
aggressiveness$isolate <- as.factor(aggressiveness$isolate)
aggressiveness$species <- as.factor(aggressiveness$species)

exp_plot <- function(y) {
  ggplot(data = aggressiveness, aes(x = experiment, y = .data[[y]])) +
    geom_boxplot() +
    theme_minimal() +
    labs(x = "", 
         y = y) +
    theme(legend.position = "none")
  }

by_exp_shoot <- exp_plot("shoot_mg")
by_exp_dry <- exp_plot("root_mg")
by_exp_length <- exp_plot("length_in")
by_exp_area <- exp_plot("area_in.2")

plot_grid(by_exp_shoot, by_exp_dry, by_exp_length, by_exp_area)
```

```{r distribution, warning=FALSE, message=FALSE}
dis_shoot <- qplot(shoot_mg, data = aggressiveness) +
  theme_minimal()
dis_dry <- qplot(root_mg, data = aggressiveness) +
    theme_minimal()
dis_length <- qplot(length_in, data = aggressiveness) +
  theme_minimal() # skew to the left, some transformation needed
dis_area <- qplot(area_in.2, data = aggressiveness) +
  theme_minimal() # also skew

plot_grid(dis_shoot, dis_dry, dis_length, dis_area)
```

```{r log transform, warning=FALSE, message=FALSE, fig.height=3, fig.width=7.5}
dis_length_log <- qplot(log(length_in), data = aggressiveness) +
  theme_minimal() 
dis_area_log <- qplot(log(area_in.2), data = aggressiveness) +
  theme_minimal() 

plot_grid(dis_length_log, dis_area_log) #improves a bit
```

## Fit mixed model

```{r fm root dry}
# not nesting isolate:species, isolates coded in a way there is no needed
# same coefficient estimates and p values without nesting
fm_root_dry <- lmer(root_mg ~ species + isolate + (1|experiment), data = aggressiveness)
resid_panel(fm_root_dry) 
anova(fm_root_dry)
emmeans(fm_root_dry, ~ species) %>% 
  cld.emmGrid()
```

```{r fm shoot dry }
fm_shoot_dry <- lmer(shoot_mg ~ species + isolate + (1|experiment), data = aggressiveness)
resid_panel(fm_shoot_dry) 
anova(fm_shoot_dry)
emmeans(fm_shoot_dry, ~ species) %>% 
  cld.emmGrid()
```

```{r fm root length }
# log transformation for normality and slight difference in variances
fm_root_length <- lmer(log(length_in) ~ species + isolate + (1|experiment), data = aggressiveness)
resid_panel(fm_root_length) # diagnostic plots look fine
anova(fm_root_length)
emmeans(fm_root_length, ~ species, type = "response") %>% # back-transform estimates
  cld.emmGrid()
```

```{r fm root area}
fm_root_area <- lmer(log(area_in.2) ~ species + isolate + (1|experiment), data = aggressiveness)
resid_panel(fm_root_area) # transformation also worked 
anova(fm_root_area)
emmeans(fm_root_area, ~ species, type = "response") %>% 
  cld.emmGrid()
```

## Differences within species (between isolates)

```{r mean separation isolates}
emmeans(fm_root_dry, ~ isolate) %>% 
  cld.emmGrid()
emmeans(fm_shoot_dry, ~ isolate) %>% 
  cld.emmGrid()
emmeans(fm_root_length, ~ isolate, type = "response") %>% 
  cld.emmGrid()
emmeans(fm_root_area, ~ isolate, type = "response") %>% 
  cld.emmGrid()
```
## Figure 4

```{r fig 4, fig.height=4.8, fig.width=3.2, warning=FALSE}
# combine the two controls for plotting
aggressiveness_plots <- aggressiveness

aggressiveness_plots$species <- str_replace_all(aggressiveness_plots$species, 
                                          c("Control Agar" = "Control", 
                                            "Control None" = "Control", 
                                            "Phytophthora sojae" = "P. sojae", 
                                            "Phytophthora sansomeana" = "P. sansomeana")
                                          )

fig_length <- aggressiveness_plots %>% 
  ggplot(aes(x = species, y = length_in, color =  species)) +
  geom_boxplot(alpha = 0.6) +
  geom_point(position = position_jitter(), alpha = 0.15, color = "gray50") +
  labs(x = "", 
         y = "Root length (in)") + 
  theme_minimal() +
  theme(legend.position = "none")  

fig_dry <- ggplot(data = aggressiveness_plots, aes(x = species, y = root_mg, color =  species)) +
  geom_boxplot(alpha = 0.8) +
  geom_point(position = position_jitter(), alpha = 0.15, color = "gray50") +
  labs(x = "", 
       y = "Root dry weigth (mg)") + 
  theme_minimal() +
  theme(legend.position = "none") 

plot_grid(fig_dry, fig_length, ncol = 1, labels = "AUTO" )
```

## Figure 5

```{r figure 5, fig.height=5, fig.width=6}
fig_5A <- aggressiveness_plots %>% 
  ggplot(aes(x = reorder (isolate, root_mg), y = root_mg, color = species)) + 
  coord_flip() +
  stat_summary(fun.data = "mean_se", aes(colour = species), alpha = 0.6) +
  geom_hline(yintercept = 20, linetype="dashed",  size = 0.7) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "",
       y = "Root dry weigth (mg)",
       color = "") 

fig_5B <- aggressiveness_plots %>% 
  filter(!is.na(length_in)) %>% 
ggplot(aes(x = reorder (isolate,length_in), y = length_in, color = species)) + 
  coord_flip() +
  stat_summary(fun.data = "mean_se", aes(colour = species), alpha = 0.8) +
  geom_hline(yintercept = 25, linetype="dashed",  size = 0.7) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "",
       y = "Root length (in)",
       color = "") 

legend <- get_legend(fig_5A + theme(legend.position = "bottom"))

plot_grid(fig_5A, fig_5B, legend, rel_heights = c(1, .1), labels = c("A", "B", ""))
```

## Save results  

```{r save agg means, message=FALSE}
m_root_dry <- emmeans(fm_root_dry, ~ isolate) %>% 
  tidy() %>% 
  mutate(measurement = "dry_root")

m_shoot_dry <- emmeans(fm_shoot_dry, ~ isolate) %>% 
  tidy() %>% 
  mutate(measurement = "dry_shoot")

m_root_length <- emmeans(fm_root_length, ~ isolate, type = "response") %>% 
  tidy() %>% 
  mutate(measurement = "root_length")

m_root_area <- emmeans(fm_root_area, ~ isolate, type = "response") %>% 
  tidy()%>% 
  mutate(measurement = "root_area")

means_aggressiveness <- bind_rows(m_root_dry, m_shoot_dry, m_root_length, m_root_area)

# estimate and response (back-transform means) in same column
back_transform_means <- means_aggressiveness %>% 
  filter(measurement %in% c("root_length", "root_area")) %>% 
  dplyr::select(-estimate) %>% 
  rename(estimate = response)

means_aggressiveness <- means_aggressiveness %>% 
  filter(!(measurement %in% c("root_length", "root_area"))) %>% 
  dplyr::select(-response)

means_aggressiveness <- bind_rows(means_aggressiveness, back_transform_means)
# write_csv(means_aggressiveness , here("output", "results", "02_aggressiveness_results-means.csv"))
```

## Correlation between aggressiveness and complexity 

```{r commbine sojae data, message=FALSE}
# import saved results
sojae_pathotype <- read_csv(here("output", "results", "02_pathotype-psojae_results.csv"))
means_aggressiveness <- read_csv(here("output", "results", "02_aggressiveness_results-means.csv"))

sojae_complexity <- sojae_pathotype %>% 
  dplyr::select(isolate_id, complexity)

sojae_aggressiveness <- means_aggressiveness %>% 
  filter(species == "Phytophthora sojae") %>% 
  dplyr::select(isolate_id = isolate, estimate, measurement) %>% 
  spread(measurement, estimate)

sojae_all_data <- left_join(sojae_complexity, sojae_aggressiveness, by = c("isolate_id")) 
```

```{r correlation}
sojae_corr <- sojae_all_data %>%
  dplyr::select(-isolate_id) %>%
  as.matrix() %>%
  rcorr(type = "pearson")
corrplot(sojae_corr$r, method = "number")
```

## Table 2

```{r table 2}
sojae_corr 
```


# Fungicide Sensitivity 

Clean data and transform data for analysis

```{r prepare fs data}
# eliminate NAs
# average colony measurements 
clean_fungicide <- fungicide %>% 
  filter(!is.na(length_a)) %>%  
  mutate(length.mm = (length_a + length_b)/2) 

# subtract the plug diameter
clean_fungicide <- clean_fungicide %>% 
  mutate(length.mm = case_when(
    length.mm > 0 ~ length.mm - 4, 
    TRUE ~ length.mm - 0
    )) %>% 
  filter(!length.mm < 0) # eliminate any negative values 

# mean growth of control plate (dose 0)
control_fungicide <- clean_fungicide %>% 
  filter(dose == 0, 
         !is.na(length.mm)) %>% 
  group_by(fungicide, experiment, isolate) %>% 
  summarise(control.mean = mean(length.mm)) 

clean_fungicide <- left_join(clean_fungicide, control_fungicide, by = c("fungicide",
                                                                        "experiment", 
                                                                        "isolate"))
# relative growth calculation
clean_fungicide <- clean_fungicide %>% 
  mutate(rel.growth = (length.mm/control.mean)*100)

# write_csv(clean_fungicide, here("output", "transform_data", "02_fungicide-sensitivity_transform.csv"))
```

Filter isolates 

```{r filter fungicide}
# check if there are isolates with no EC50
# isolates with >50% rel growth in the higher dose 
no_EC50 <- clean_fungicide %>% 
  group_by(fungicide, experiment, isolate, dose) %>% 
  summarise(rel.growth = mean(rel.growth)) %>% 
  filter(fungicide %in% c("Mefenoxam",
                          "Metalaxyl", 
                          "Ethaboxam") & dose == 1 & rel.growth > 50, 
         fungicide == "Azoxystrobin" & dose == 10 & rel.growth > 50)

# check if there are isolates with hormetic effect
# >100% rel growth at lower dose
hormetic_isolates <- clean_fungicide %>% 
  group_by(fungicide, experiment, isolate, dose) %>% 
  summarise(rel.growth = mean(rel.growth)) %>% 
  filter(rel.growth > 100, 
         !dose == 0) %>% 
  dplyr::select(fungicide, experiment, isolate)

hormetic_isolates %>% 
  group_by(fungicide, experiment, isolate) %>% 
  n_distinct()

hormetic_data <- semi_join(clean_fungicide, hormetic_isolates, by = c("fungicide",
                                                                        "experiment", 
                                                                        "isolate"))
```

```{r fit log logistic models, warning=FALSE }
# fit log-logistic models with 3 and 4 parameters 
nest_fungicide <- clean_fungicide %>% 
  group_by(fungicide, experiment, isolate) %>% 
  nest()

fit_LL3 <- function(dataset){
  drm(rel.growth ~ dose, fct = LL.3(names = c("slope", "upper", "EC50")), data = dataset)
}

fit_LL4 <- function(dataset){
  drm(rel.growth ~ dose, fct = LL.4(names = c("slope", "lower", "upper", "EC50")), data = dataset)
}

log_models <- nest_fungicide %>% 
  mutate(ll3.model = purrr::map(data, fit_LL3), 
         ll4.model = purrr::map(data, fit_LL4), 
         ll3.fit = purrr::map(ll3.model, mselect), 
         ll4.fit = purrr::map(ll4.model, mselect) 
         )
# mselect did not work inside function, use it individually to extract model fit info
```

```{r select best model}
log_selection <- log_models %>% 
  dplyr::select(ll3.fit, ll4.fit) %>% 
  gather(ll3.fit, ll4.fit, key = "model", value = "model.fit") %>%   
  unnest_wider(model.fit)

# select based on lower AIC
log_selection %>% 
  dplyr::select(fungicide, isolate, model, IC) %>% 
  spread(model, IC) %>% 
  filter(ll3.fit < ll4.fit) %>% 
  n_distinct() 
```
From the 256 models fit (experiment, fungicide and isolation combination), only 66 had higher AIC value for 4 parameters model (LL.4)

```{r fit p values}
# lack of fit p-values 
log_selection %>% 
  filter(`Lack of fit` > 0.05) %>% 
  group_by(model)%>% 
  count() %>% 
  knitr::kable()
```

More non-significant p-values for lack of fit  for the 4 parameters model (LL.3).  
Overall, LL.4 model better fit for most isolates

```{r EC50s, message=FALSE}
# extract absolute EC50s from LL.4 model
abs_EC50 <- function(models){
  ED(models, respLev = c(50), type = "absolute", interval = "delta")
}

log_models <- log_models %>% 
  mutate(ll4.EC50 = purrr::map(ll4.model, abs_EC50))

log_EC50 <- log_models %>% 
  dplyr::select(ll4.EC50) %>% 
  unnest_wider(ll4.EC50) %>% 
  rename(EC50 = ...1  , 
         std.error = ...2 ,
         lower = ...3, 
         upper = ...4)
```

Fit hormetic models to data with hormetic effects.

```{r fit hormetic models, message=FALSE, warning=FALSE}
nest_hormetic <- hormetic_data %>% 
  group_by(fungicide, experiment, isolate) %>% 
  nest()

fit_CRS <- function(dataset){
  drm(rel.growth ~ dose, fct = CRS.4c(), data = dataset)
}

fit_BC4 <- function(dataset){
  drm(rel.growth ~ dose, fct = BC.4(), data = dataset)
}

# fit hormetic models and compared it to LL.4
hormetic_models <- nest_hormetic %>% 
  mutate(crs.model = purrr::map(data, fit_CRS),
         bc4.model = purrr::map(data, fit_BC4 ),
         crs.fit = purrr::map(crs.model, mselect), 
         bc4.fit = purrr::map(bc4.model, mselect)
         )
```

```{r hormetic model selection}
hormetic_selection <- hormetic_models %>% 
  dplyr::select(crs.fit, bc4.fit)

hormetic_selection <- hormetic_selection %>% 
  gather(crs.fit, bc4.fit,  key = "model", value = "model.fit") %>%   
  unnest_wider(model.fit)

# add LL4 fit 
hormetic_selection <- semi_join(log_selection, hormetic_isolates) %>% 
  filter(model == "ll4.fit") %>% 
  bind_rows(hormetic_selection) %>% 
  arrange(isolate, IC)

# best hormetic model
hormetic_selection %>% 
  dplyr::select(fungicide, experiment, isolate, model, IC) %>% 
  spread(model, IC) %>% 
  filter(crs.fit < bc4.fit) %>% 
  n_distinct()
  

```

111 hormetic models fit, only 13 had higher for BC.4

```{r lack of fit hormetic}
# compare hormetic models vs LL.4
hormetic_selection %>% 
  dplyr::select(fungicide, experiment, isolate, model, IC) %>% 
  spread(model, IC) %>% 
  filter(ll4.fit > bc4.fit) %>% 
  n_distinct()
  
hormetic_selection %>% 
  filter(`Lack of fit` > 0.05) %>% 
  group_by(model) %>% 
  count() %>% 
  knitr::kable()
```

overall bc4 better fit based on lower AIC
no lack of fit for ll4 for most isolates with hormetic effect


```{r both data}
# check if 95 CI of exp 1 includes EC50 of exp 2
log_EC50$experiment <- as.factor(log_EC50$experiment)
log_EC50$isolate <- as.factor(log_EC50$isolate)

log_EC50 %>% 
  filter(!isolate == "13") %>% # 95% CI too large for plotting
  ggplot(aes(x = isolate, y = EC50, color = experiment)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                  position = position_jitterdodge()) +
  facet_wrap(~ fungicide, scales = "free_y") +
  theme_minimal() +
  panel_border()
```

Overall most EC50 of second experiment fall into the 95% CI of EC50 first. Experiment is repoducible so combine the data of both experiments. 

```{r LL.4 combine data, message=FALSE, warning=FALSE}
# fit LL.4 model with pooled data
nest_pooled <- clean_fungicide %>% 
  group_by(fungicide, isolate) %>% 
  nest() 

log_pooled <- nest_pooled %>% 
  mutate(ll4.model = purrr::map(data, fit_LL4),
         ll4.fit = purrr::map(ll4.model, mselect), 
         abs.EC50 = purrr::map(ll4.model, abs_EC50)
         ) 

pooled_EC50 <- log_pooled %>% 
  dplyr::select(abs.EC50) %>% 
  unnest_wider(abs.EC50) %>% 
  rename(EC50 = ...1  , 
         std_error = ...2 ,
         lower = ...3, 
         upper = ...4) 
```

```{r pooled hormetic model, message=FALSE, warning=FALSE}
# fit pool hormetic model
nest_hormetic_pooled <- hormetic_data %>% 
  group_by(fungicide, isolate) %>% 
  nest() 

hormetic_pooled <- nest_hormetic_pooled %>% 
  filter(!fungicide == "ethaboxam") # not working abs_Ec50 come back

hormetic_pooled <- hormetic_pooled %>% 
  mutate(bc4.model = purrr::map(data, fit_BC4),
         bc4.fit = purrr::map(bc4.model, mselect),
         abs.EC50 = purrr::map(bc4.model, abs_EC50)
         )

hormetic_EC50 <- hormetic_pooled %>% 
  dplyr::select(abs.EC50) %>% 
  unnest_wider(abs.EC50) %>% 
  rename(EC50 = ...1  , 
         std.error = ...2 ,
         lower = ...3, 
         upper = ...4) %>% 
  mutate(model = "BC4")
```

```{r compare EC50, fig.height=3, fig.width=8}
# compare hormetic EC50 with LL4 
compare_EC50 <- semi_join(pooled_EC50, hormetic_isolates, by = c("fungicide", "isolate")) %>% 
  mutate(model = "LL4") 

compare_EC50 <- bind_rows(compare_EC50, hormetic_EC50)


compare_EC50$isolate <- as.factor(compare_EC50$isolate)

compare_EC50 %>% 
  filter(!fungicide == "ethaboxam") %>% # problem calculating EC50
  ggplot(aes(x = isolate, y = EC50, color = model)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                  position = position_jitterdodge()) +
  facet_wrap(~ fungicide, scales = "free") +
  theme_minimal() +
  panel_border()
```

Lower EC50 for the hormetic models.

## Figure 6

```{r figure 6, fig.height=5, fig.width=5.5}
EC50_plot <- pooled_EC50 %>% 
  mutate(species = case_when(
    isolate %in% c(1, 5, 7:10, 14:16) ~ "P. sojae",
    TRUE ~ "P. sansomeana"
  ))

EC50_plot$fungicide <- str_replace_all(EC50_plot$fungicide, 
                                       c("azoxystrobin" = "Azoxystrobin", 
                                         "ethaboxam" = "Ethaboxam", 
                                         "mefenoxam" = "Mefenoxam", 
                                         "metalaxyl" = "Metalaxyl")
                                       )

ggplot(EC50_plot, aes(x = EC50, fill = species)) +
  geom_histogram(alpha = 0.6, color = "white", bins = 10) +
  facet_wrap(~ fungicide, scales = "free_x") +
  labs(x = expression('EC'[50]*' (μg ml'^-1*')'), 
       y = "Frequency", 
       fill = "") +
  theme_minimal() +
  panel_border() +
  theme(legend.position = "bottom", 
        strip.text = element_text(face = "bold"), 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        panel.grid.minor.y = element_blank()) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) 
```
```{r save EC50}
# write_csv(pooled_EC50, here("output", "results", "02_fungicide-sensitivity_EC50.csv"))
# write_csv(compare_EC50, here("output", "results", "02_fungicide-sensitivity_EC50_hormetic-isolates.csv"))
```

